---
title: "STEPS Ecuador 2025 HTA Data Audit"
subtitle: "Initial reproducible audit; predictive modeling is out of scope"
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
source("../R/00_config.R")
source("../R/01_validate_inputs.R")
source("../R/02_derive_hta.R")
source("../R/03_audit_tables.R")

ecu_raw <- hta_read_ecu_data()
hta_check_packages("knitr")
validation <- hta_assert_valid_inputs(ecu_raw)
ecu_audit <- hta_derive_preliminary(ecu_raw)
```

## Privacy and Scope

This report reads `ECU_data.xlsx` from a configurable local path. Raw microdata
are not committed to the public repository and no derived `.rds` artifacts are
created in this phase.

The audit is limited to schema validation, blood pressure completeness, and a
preliminary hypertension derivation. Predictive modeling is explicitly out of
scope for this phase.

## Source Data

The workbook path is:

```{r}
normalizePath(hta_config$data_path, winslash = "/", mustWork = FALSE)
```

## Input Validation

### Dimensions

```{r}
knitr::kable(validation$dimensions)
```

### Required Variables

```{r}
knitr::kable(validation$variables)
```

Required variables include survey design fields, three systolic and diastolic
blood pressure readings, history of hypertension diagnosis, and medication
status.

## Blood Pressure Completeness

```{r}
knitr::kable(hta_bp_completeness_table(hta_clean_bp_readings(ecu_raw)))
```

```{r}
knitr::kable(hta_bp_reading_count_table(ecu_audit))
```

## Preliminary HTA Derivation

Preliminary SBP and DBP are computed as row means across valid readings:

- SBP: `m4a`, `m5a`, `m6a`
- DBP: `m4b`, `m5b`, `m6b`

The preliminary `hta_140_90_or_meds` flag is positive when mean SBP is at least
140 mmHg, mean DBP is at least 90 mmHg, or `m7 == 1` indicates current
medication for hypertension.

```{r}
knitr::kable(hta_preliminary_summary_table(ecu_audit))
```

## Methodological Limitations

- This is a preliminary audit derivation, not the final epidemiologic analysis.
- Missing-value handling is limited to common STEPS special codes and should be
  reconciled with the official WHO/STEPS processing pipeline before publication.
- The preliminary blood pressure mean uses all available valid readings. Final
  reporting may require a protocol-specific average, such as the average of the
  second and third readings when both exist.
- The current flag includes measured blood pressure and medication. It should not
  be used as a predictive-model target with the same blood pressure readings as
  predictors, because that would introduce information leakage.
- Survey-weighted estimates and confidence intervals are not produced in this
  initial audit report.

## Modeling Status

No predictive models are trained in this phase. No model objects, `.rds` files,
or derived microdata are saved by this report.
